---
title: "EDLD610_FinalProj"
author: "Andrew Fridman"
date: "February 27, 2019"
output: html_document
---

```{r Setup, include = FALSE}
#Load libraries
library(colorblindr)
library(dplyr)
library(here)
library(ggplot2)
library(gsubfn)
library(lubridate)
library(psych)
library(rio)
library(snakecase)
library(tidyverse)
```

```{r Import data}
#Import arabica data
arabica <- rio::import("arabica_data_cleaned.csv")
colnames(arabica) <- tolower(colnames(arabica))
#View(arabica)

#Import robusta data
robusta <- rio::import("robusta_data_cleaned.csv") 
colnames(robusta) <- tolower(colnames(robusta))
#View(robusta)
```

```{r Clean-up data}
#Tidy arabica data
arabica_t <- arabica %>% 
                     select(species, country.of.origin, aroma, flavor, aftertaste,       
                            acidity, body, balance, uniformity, clean.cup, sweetness, cupper.points, 
                            total.cup.points) %>%
                     rename(country = "country.of.origin",
                            clean = "clean.cup",
                            cupper_points = "cupper.points",
                            total_points = "total.cup.points") %>% 
                     filter(total_points > 0 & 
                            country != "") %>% 
                     mutate(country = as.factor(country),
                            species = as.factor(species)) %>% 
                     mutate(country = fct_recode(country, "Ivory Coast" = "Cote d?Ivoire"),
                            country = fct_recode(country, "Tanzania" = "Tanzania, United Republic Of"))
  
#Tidy arabica data
robusta_t <- robusta %>% 
                     select(species, country.of.origin, fragrance...aroma, flavor,
                            aftertaste, salt...acid, mouthfeel, balance, uniform.cup, clean.cup, 
                            bitter...sweet, cupper.points, total.cup.points) %>% 
                       rename(country = "country.of.origin",
                              aroma = "fragrance...aroma",
                              acidity = "salt...acid",
                              sweetness = "bitter...sweet",
                              body = "mouthfeel",
                              uniformity = "uniform.cup",
                              clean = "clean.cup",
                              cupper_points = "cupper.points",
                              total_points = "total.cup.points") %>% 
                     filter(total_points > 0 & 
                            country != "") %>% 
                     mutate(country = as.factor(country),
                            species = as.factor(species))

#Join databases
coffee <- rbind(arabica_t, robusta_t)

#Create regions
regions <- data.frame(country = c("Brazil", "Burundi", "China", "Colombia", "Costa Rica", "Ivory Coast",
                                  "Ecuador", "El Salvador", "Ethiopia", "Guatemala", "Haiti",
                                  "Honduras", "India", "Indonesia", "Japan", "Kenya", "Laos", "Malawi",
                                  "Mauritius", "Mexico", "Myanmar", "Nicaragua", "Panama", "Papua New Guinea", 
                                  "Peru", "Philippines", "Rwanda", "Taiwan", "Tanzania", "Thailand", "Uganda", 
                                  "United States", "United States (Hawaii)", "United States (Puerto Rico)", 
                                  "Vietnam", "Zambia"),  
                      region = c("South America", "Africa", "Asia", "South America", "Central America",    
                                 "Africa", "South America", "Central America", "Africa", "Central America",                                       "Central America", "Central America", "Asia", "Southeast Asia", "Asia", 
                                 "Africa", "Southeast Asia", "Africa", "Africa", "North America", 
                                 "Southeast Asia", "Central America", "Central America", "Oceania", 
                                 "South America", "Southeast Asia", "Africa", "Asia", "Africa", "Southeast Asia",
                                 "Africa", "North America", "North America", "North America", "Southeast Asia", 
                                 "Africa"))

#Join databases
coffee_join <- left_join(coffee, regions)
```

##TN: 
Great data wrangling! 

```{r Visualization 1 - Total reviews by country and region}
#Tidy data
coffee_reviews <- coffee_join %>% 
                              select(country, region) %>% 
                              count(country, region)

#Histogram: Number of reviews by country (with region highlighted)
ggplot(coffee_reviews, aes(fct_reorder(country, n), n)) +
       geom_col(aes(fill = region),
                color = "white",
                alpha = 0.7) +
       coord_flip() +
       scale_fill_OkabeIto() +
       theme_minimal() +
       labs(title = "Total Reviews by Country",
            x = "Country",
            y = "Reviews")

#Tidy data
coffee_reviews_reg <- coffee_join %>% 
                                  select(region) %>% 
                                  count(region)

#Histogram: Number of reviews by country
ggplot(coffee_reviews_reg, aes(fct_reorder(region, n), n)) +
       geom_col(aes(fill = region),
                color = "white",
                alpha = 0.7) +
       coord_flip() +
       scale_fill_OkabeIto() +
       theme_minimal() +
       labs(title = "Total Reviews by Region",
            x = "Region",
            y = "Reviews")

```

##TN:

I wish there was a little description on what you goal for each visualization is. 
These graphs look so pretty! 

Are you going to try to show both? or are you still trying to decide on one of them?
Unless you care to show something specific of certain countries, if youre trying to choose one, I feel that the graph that is just divided by world regions is clearer. You can understand what world regions have more reviews , and that seems to be the main message you want to transmit (unless I'm wrong!) , while the grah by country can be a bit more ocnfusing because the country with most reviews does not correspond to the world region with more reviews for example - un less that is the point you want to transmit with this visualization! 

Great use of the okabe_ito color palette btw! :) 


```{r Visualization 2 - Total points by country and region}
#Tidy data
coffee_countries <- coffee_join %>% 
                       select(country, region, total_points) %>% 
                       group_by(country, region) %>% 
                       summarise(total_points_m = mean(total_points, na.rm = TRUE)) 

#Total score by country (with regions highlighted)
ggplot(coffee_countries, aes(fct_reorder(country, total_points_m), total_points_m)) +
       geom_col(aes(fill = region),
                color = "white",
                alpha = 0.7) +
       geom_text(aes(country, total_points_m, label = round(total_points_m, 2)),
                 nudge_y = -2,
                 size = 2.5) +
       coord_flip() +
       scale_fill_OkabeIto() +
       theme_minimal() +
       labs(title = "Total Points by Country",
            x = "Country",
            y = "Total Points")

#Tidy data
coffee_regions <- coffee_join %>%
                              select(region, total_points) %>% 
                              group_by(region) %>% 
                              summarise(total_points_m = mean(total_points, na.rm = TRUE))

#Total score by region
ggplot(coffee_regions, aes(fct_reorder(region, total_points_m), total_points_m)) +
       geom_col(aes(fill = region),
                color = "white",
                alpha = 0.7) +
       geom_text(aes(region, total_points_m, label = round(total_points_m, 2)),
                 nudge_y = -2,
                 size = 2.5) +
       coord_flip() +
       scale_fill_OkabeIto() +
       theme_minimal() +
       labs(title = "Total Points by Region",
            x = "Region",
            y = "Total Points")
```

```{r Visualization 3 - Mean quality metrics by species}
#Tidy data
coffee_species <- coffee %>% 
                         select(-country) %>% #I took away "-year" because year is already not part of the data frame, it was giving me an error otherwise
                         group_by(species) %>% 
                         summarise(Aroma = mean(aroma, na.rm = TRUE),
                                   Flavor = mean(flavor, na.rm = TRUE),
                                   Aftertaste = mean(aftertaste, na.rm = TRUE),
                                   Acidity = mean(acidity, na.rm = TRUE),
                                   Body = mean(body, na.rm = TRUE),
                                   Balance = mean(balance, na.rm = TRUE),
                                   Uniformity = mean(uniformity, na.rm = TRUE),
                                   Clean = mean(clean, na.rm = TRUE),
                                   Sweetness = mean(sweetness, na.rm = TRUE)) %>% 
                         gather(metric, score, -1) %>% 
                         mutate(metric = as.factor(metric))

#Mean quality metrics by species
ggplot(coffee_species, aes(fct_reorder(metric, score), score)) +
       geom_col(aes(fill = species), 
                position = "dodge",
                alpha = 0.7) +
       coord_flip() +
       scale_color_OkabeIto() +
       theme_minimal() +
       theme_minimal() +
       labs(title = "Comparison of quality metric scores by coffee bean species",
            x = "Quality Metrics",
            y = "Score")

######Other possible versions#####

#Line difference in mean quality metrics by species
ggplot(coffee_species, aes(fct_reorder(metric, score), score)) +
       geom_point(aes(color = species),
                  alpha = 0.7,
                  size = 5) +
       geom_line(aes(group = metric)) +
       coord_flip() +
       scale_color_OkabeIto() +
       theme_minimal() +
       labs(title = "Comparison of quality metric scores by coffee bean species",
            x = "Quality Metrics",
            y = "Score")

#Tidy data
coffee_species_diff <- coffee_species %>% 
                                      spread(species, score) %>% 
                                      ungroup() %>% 
                                      mutate(diff = Arabica - Robusta,
                                             metric = fct_reorder(metric, diff))

#Point difference in mean quality metrics by species
ggplot(coffee_species_diff, aes(fct_reorder(metric, diff), diff)) +
       geom_hline(yintercept = 0,
                  color = "gray70",
                  size = 1) +
       geom_point(color = "cornflowerblue",
                  alpha = 0.7,
                  size = 3) +
       geom_line(aes(group = metric)) +
       coord_flip() +
       theme_minimal() +
       labs(title = "Difference in quality metric scores by coffee bean species (Arabica > Robusta)",
            x = "Quality Metrics",
            y = "Difference in Score (Points)")
```

##TN:

Nice plot ideas!! if I get the message you are trying to trasnmit here (that the quality aspect that robusta and arabica differ the most is Sweetness, and aftertaste in the second place, but by way less of a difference) - I would say that the second version of the plot is my favorite! , I believe is the one transmitting the point the best and it's easy to get! The bar plot is a bit confusing, too many bars stacked together and is harder to identify sweetness as the one with the greates difference. A similar thing happens with the third version of the plot, it's easier to see that sweetnes has the greatest difference but it is harder to interpret. 

